NGO Repression as a Predictor of Worsening Human Rights Abuses

### 2021-11-19 manuscript/appendix.Rmd
Final minor edits pre-submission

---
title: 'Appendix for "NGO Repression as a Predictor of Worsening Human Rights Abuses"'
short-title: 'Appendix for "NGO Repression"'
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "November 19, 2021"
# published: Under review
code-repo: "Access the code and full analysis notebook at <https://stats.andrewheiss.com/canary-ngos/>"
reference-section-title: References
bibliography: bibliography.bib
toc: false
appendix: true
mainfont: Cochineal
sansfont: Inter
editor_options: 
  chunk_output_type: console
---

@@

# Knitr options
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  tidy.opts = list(width.cutoff = 120),  # Code width
  fig.retina = 3, dpi = conditional_dpi,
  fig.width = 7, fig.asp = 0.618,
  fig.align = conditional_align, out.width = "100%",
  fig.path = "output/figures/",
  cache.path = "output/_cache/",
  fig.process = function(x) {  # Remove "-1" from figure names
    x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
    if (file.rename(x, x2)) x2 else x
  }
)

@@

We use Stan [@stan] through R [@r-project] and brms [@brms-jss] to estimate our models. We generate 4 MCMC chains for each model with 2,000 iterations in each chain, 1,000 of which are used for warmup. All chains converge; we assess convergence with visual inspection.

@@

$$
\begin{aligned}
& \textbf{Likelihood for ordinal outcome} \\
& \text{Political Terror Scores (PTS)} \\
y_{i, t+1} \sim&\ \operatorname{Categorical}(p_{i, t + 1}) \\
\ \\
& \textbf{Parameters} \\
\operatorname{logit}(p_i) =&\ \alpha_{j[i]} - \phi_{k, t + 1} \\
\phi_{k, t + 1} =&\ \beta_1 \text{PTS}_{i, t} + \beta_2 \text{PTS}_{i, t - 1} + \\
&\ \beta_3 \text{Outcome}_{i, t} + \beta_4 \text{Polyarchy}_{i, t} + \\
&\ \beta_5 \log (\text{GDP per capita}) + \\
&\ \beta_6 \text{Trade as } \% \text{ of GDP}_{i, t} + \\
&\ \beta_7 \text{Armed conflict}_{i, t} \\ \\
\alpha_j \sim&\ \mathcal{N} (\mu_{\alpha_j}, \sigma_{\alpha_j}), \text{ for country } j \text{ in } 1 .. J & \text{[country-specific intercepts]} \\ \\
\ \\
& \textbf{Priors} \\
\mu_{\alpha_j} \sim&\ \mathcal{N}(0, 10) & \text{[country-specific intercepts]} \\
\beta_{1-7} \sim&\ \mathcal{N}(0, 3) & \text{[population effects]} \\
\sigma_{\alpha_j} \sim&\ \operatorname{Cauchy}(0, 1) & \text{[sd for population \& country]}
\end{aligned}
$$

@@

```{r echo=TRUE, eval=FALSE}
# Ordinal logistic regression for political terror
# (replace `cs_repression` with different measures of de jure and 
# de facto civil society repression: barriers_total, advocacy, entry, 
# funding, and v2csreprss)
brm(
  bf(PTS_lead1 ~ cs_repression + cs_repression_lag1 +
       PTS + v2x_polyarchy + gdpcap_log +
       un_trade_pct_gdp + armed_conflict + (1 | gwcode)),
  family = cumulative(),
  prior = c(set_prior("normal(0, 10)", class = "Intercept"),
            set_prior("normal(0, 3)", class = "b"),
            set_prior("cauchy(0, 1)", class = "sd")),
  ...
)
```

>>>>

---
title: 'Online appendix for "NGO Repression as a Predictor of Worsening Human Rights Abuses"'
short-title: 'Online appendix for "NGO Repression"'
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "November 19, 2021"
# published: Under review
code-repo: "Access the code and full analysis notebook at <https://stats.andrewheiss.com/cautioning-canary/>"
reference-section-title: References
bibliography: bibliography.bib
appendix: true
toc: true
title-page: false
mainfont: Cochineal
sansfont: Inter
editor_options: 
  chunk_output_type: console
---

@@

# Knitr options
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  tidy.opts = list(width.cutoff = 120),  # Code width
  fig.retina = 3, dpi = conditional_dpi,
  fig.width = 7, fig.asp = 0.618,
  fig.align = conditional_align, out.width = "100%",
  fig.path = "output/figures-appendix/",
  cache.path = "output/_cache/",
  fig.process = function(x) {  # Remove "-1" from figure names
    x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
    if (file.rename(x, x2)) x2 else x
  }
)

@@

We use Stan [@stan] through R [@r-project] and **brms** [@brms-jss] to estimate our models. We generate 4 MCMC chains for each model with 2,000 iterations in each chain, 1,000 of which are used for warmup. All chains converge; we assess convergence with visual inspection.

@@

$$
\begin{aligned}
& \textbf{Likelihood for ordinal outcome} \\
& \text{Political Terror Scores (PTS)} \\
y_{i, t+1} \sim&\ \operatorname{Categorical}(p_{i, t + 1}) \\
\ \\
& \textbf{Parameters} \\
\operatorname{logit}(p_i) =&\ \alpha_{j[i]} - \phi_{k, t + 1} \\
\phi_{k, t + 1} =&\ \beta_1 \text{PTS}_{i, t} + \beta_2 \text{PTS}_{i, t - 1} + \\
&\ \beta_3 \text{Outcome}_{i, t} + \beta_4 \text{Polyarchy}_{i, t} + \\
&\ \beta_5 \log (\text{GDP per capita}) + \\
&\ \beta_6 \text{Trade as } \% \text{ of GDP}_{i, t} + \\
&\ \beta_7 \text{Armed conflict}_{i, t} \\ \\
\alpha_j \sim&\ \mathcal{N} (\mu_{\alpha_j}, \sigma_{\alpha_j}), \text{ for country } j \text{ in } 1 .. J & \text{[country-specific intercepts]} \\ \\
\ \\
& \textbf{Priors} \\
\mu_{\alpha_j} \sim&\ \mathcal{N}(0, 3) & \text{[country-specific intercepts]} \\
\beta_{1-7} \sim&\ \mathcal{N}(0, 3) & \text{[population effects]} \\
\sigma_{\alpha_j} \sim&\ \operatorname{Cauchy}(0, 1) & \text{[sd for population \& country]}
\end{aligned}
$$

@@

```{r echo=TRUE, eval=FALSE}
# Ordinal logistic regression for political terror
# (replace `cs_repression` with different measures of de jure and 
# de facto civil society repression: barriers_total, advocacy, entry, 
# funding, and v2csreprss)
brm(
  bf(PTS_lead1 ~ cs_repression + cs_repression_lag1 +
       PTS + v2x_polyarchy + gdpcap_log +
       un_trade_pct_gdp + armed_conflict + (1 | gwcode)),
  family = cumulative(),
  prior = c(set_prior("normal(0, 3)", class = "Intercept"),
            set_prior("normal(0, 3)", class = "b"),
            set_prior("cauchy(0, 1)", class = "sd")),
  ...
)
```

### 2021-11-19 manuscript/manuscript.Rmd
Final minor edits pre-submission

```{r plot-mfx-e1b, fig.width=7.5, fig.height=4, fig.asp=NULL, fig.cap="Marginal effects of increasing anti-NGO legal barriers on latent human rights scoress"}
out_mfx_e1b <- mfx_e1b_lhr %>% 
  mutate(x_rounded = floor(effect1__)) %>% 
  group_by(x_rounded, plot_var_nice) %>% 
  slice(1) %>% 
  ungroup() %>% 
  mutate(plot_var_nice = fct_relabel(
    plot_var_nice, ~str_to_sentence(str_remove(., "Barriers to ")))) %>% 
  ggplot(aes(x = x_rounded, y = estimate__, color = plot_var_nice)) +
  geom_line(size = 0.5, color = "grey90") +
  geom_pointrange(aes(ymin = lower__, ymax = upper__),
                  fatten = 1, size = 1) +
  scale_x_continuous(breaks = 0:9) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  guides(color = "none") +
  labs(x = "Count of legal barriers",
       y = "Predicted latent human rights value",
       title = "Latent human rights",
       caption = "(Higher values = better protection)") +
  facet_grid(cols = vars(plot_var_nice), scales = "free_x", space = "free_x") +
  theme_ngo() +
  theme(panel.grid.major.x = element_blank())

@@

```{r tbl-rmse}
# These numbers also come from the analysis notebook at 
# https://stats.andrewheiss.com/cautioning-canary/04_predictions.html
baseline_rmse <- 0.1862156
tbl_rmse <- tribble(
  ~`Latent human rights values`, ~RMSE,
                  "Baseline",    baseline_rmse,
      "Total legal barriers",    0.1872181,
      "Barriers to advocacy",    0.1872011,
         "Barriers to entry",    0.1862635,
       "Barriers to funding",    0.1879768,
  "Civil society repression",    0.1857038
) %>% 
  mutate(`% change from baseline` = (RMSE - baseline_rmse) / baseline_rmse,
         `% change from baseline` = percent(`% change from baseline`, accuracy = 0.01)) %>% 
  mutate(`% change from baseline` = recode(`% change from baseline`, "0.00%" = "—")) %>% 
  mutate(RMSE = round(RMSE, 4))

>>>>

```{r plot-mfx-e1b, fig.width=7.5, fig.height=4, fig.asp=NULL, fig.cap="Marginal effects of increasing anti-NGO legal barriers on latent human rights values"}
out_mfx_e1b <- mfx_e1b_lhr %>% 
  mutate(x_rounded = floor(effect1__)) %>% 
  group_by(x_rounded, plot_var_nice) %>% 
  slice(1) %>% 
  ungroup() %>% 
  mutate(plot_var_nice = fct_relabel(
    plot_var_nice, ~str_to_sentence(str_remove(., "Barriers to ")))) %>% 
  ggplot(aes(x = x_rounded, y = estimate__, color = plot_var_nice)) +
  geom_line(size = 0.5, color = "grey90") +
  geom_pointrange(aes(ymin = lower__, ymax = upper__),
                  fatten = 1, size = 1) +
  scale_x_continuous(breaks = 0:9) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  guides(color = "none") +
  labs(x = "Count of legal barriers",
       y = "Predicted latent human rights value",
       title = "Latent human rights",
       caption = "(Higher values = better protection)") +
  facet_grid(cols = vars(plot_var_nice), scales = "free_x", space = "free_x") +
  theme_ngo() +
  theme(panel.grid.major.x = element_blank())

@@

```{r tbl-rmse}
# These numbers also come from the analysis notebook at 
# https://stats.andrewheiss.com/cautioning-canary/04_predictions.html
baseline_rmse <- 0.1862156
tbl_rmse <- tribble(
  ~`Latent human rights values`, ~RMSE,
            "Baseline model",    baseline_rmse,
      "Total legal barriers",    0.1872181,
      "Barriers to advocacy",    0.1872011,
         "Barriers to entry",    0.1862635,
       "Barriers to funding",    0.1879768,
  "Civil society repression",    0.1857038
) %>% 
  mutate(`% change from baseline` = (RMSE - baseline_rmse) / baseline_rmse,
         `% change from baseline` = percent(`% change from baseline`, accuracy = 0.01)) %>% 
  mutate(`% change from baseline` = recode(`% change from baseline`, "0.00%" = "—")) %>% 
  mutate(RMSE = round(RMSE, 4))
